Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-05 paper/paper.tex
fix: sentinel hypothesis

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain's baseline realizes
a ``radar'' function to
detect subjectively salient and unexpected events that
are the most important for unfolding behavior,
including resources,
the securing of mates, and protection.
%
In fact,
environmental stimuli important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
More specifically,
according to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
Highly efficient neurobiological processing of social cues is exemplified by
facial judgments being routinely processed in less than 100 ms
\citep{bar2006very}.
Our human ancestors were probably among the few organisms that
were more likely to be killed by members of their own species
rather than predators from another species.
Being able to detect the intention and upcoming actions of conspecifics
turned into a decisive evolutionary advantage
\citep{frith2010social}.
%
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and
the complex relationships between them
\citep{schilbach2008minds}.

\textbf{XXX Elvis: Your focus in this subsection ought to be on the SC itself not POMPDs. -> agree I will fix it later.}
 \begin{itemize}
    \item Processing social information has been proposed to underlie the
    physiological baseline of human brain function
    \citep{schilbach2008minds}. This was later challenged by observing
    analogues of the DMN in monkeys \citep{mantini2011default}
    and rats \citep{lu2012rat}, with
    supposedly less advanced social capacities.
    The MDPs parsimoneously explain the dominance of social content in
    human mental scenes by their extremely high relevance to humans
    reflected by high values for social information in the value matrix
    \citep{baker2009action}.
    \item How can a same neurobiological circuit be equally important
    for baseline house-keeping functions and specific task performance?
    Our computational account can explain why the DMN is implicated
    in both a goal-directed task and an idlying rest cognitive set,
    if environmental relevance is processed, manipulated, and retained
    also as a baseline function of human mental activity.
    In tasks, the policy and value matrices are updated to optimize short-term action,
    whereas, at rest, these parameter updates may
    improve especially mid- and long-term action.
    \item The sentinel account does not provide a formal mechanisms of
    how attention and memory resources are exactly reallocated when
    encountering a salient environmental stimulus. \textbf{XXX remove the next sentence. Danilo: can you help me to correct the sentence instead.
    There appears to be a cool link between cognition and formal theory
    here. Elvis: I'd've helped you correct it but, i don't quite understand what you wanted to say. For ex, note that $\sum_{a \in \mathcal A}\pi(a|s) = 1$ for any state $s$,
  and so it doesn't quite make sense to talk about weighting things with ``$\sum_{a \in \mathcal A}\pi(a|s)$''.}
    In contrast,
    the Bellman recursion is weighted at each step by
    $\sum_{a \in A}\pi(a|s)$. This factor weighs which action cascades
    are recursively explored by the agent and which decision trees are neglected.
    This directly implies allocation of attention and computational load.
    \item In sum,
    MDPs imply a ``radar'' function of monitoring the environment for salient information
    in general and explain social thinking as a physiological brain baseline
    as an important special case, without excluding other topics
    contemplated in continuous mental activity.
 \end{itemize}

>>>>

\subsection{Sentinel Hypothesis}
Processing self-relevant information was perhaps the first
cognitive account that was proposed for the DMN \citep{gusnard2001medial}.
Since then,
different investigators have speculated that neural activity in the DMN
may reflect the brain’s relentless tracking of
relevance in the environment
as an advantageous evolutionary adaptation.
According to this cognitive account, the brain\'s baseline realizes
a ``radar'' function to
detect subjectively salient object and unexpected events in the environment.
The sentinel account does however not provide a formal mechanisms of
how attention and memory resources are exactly reallocated when
encountering a salient environmental stimulus.
In our formal account,
action cascades
are recursively explored by human agents. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to get explored.
This mechanism, mediated by the Bellman equation,
directly implies allocation of attention and computational load
to relevant objects and events in the environment.
%
Further,
our computational account provides an explanation for the known implication
of the DMN
in both certain goal-directed experimental tasks and
especially in the idlying rest cognitive set.
Both cognitive sets may entail DMN recruitment
if environmental relevance is processed, manipulated, and used for prediction.
In tasks, the policy and value matrices are updated to optimize short-term action.
At rest, these parameter updates may
improve especially mid- and long-term action.
We thus provide answers to the unsettled question why the involvement
of the same neurobiological brain circuit, the DMN, has been documented
for baseline house-keeping functions and specific task performance.

In fact,
environmental stimuli especially important for humans are frequently of
social nature. This is probably unsurprising
given that a key property of the human species is
the complexity of their social systems
\citep{tomasello2009cultural}.
According to the ``social brain hypothesis'',
the human brain did not evolve to solve problems of the
physical environment, but to form and maintain increasingly complex
social systems to solve ecological problems by means of social relationships
\citep{whiten1988machiavellian}.
This may explain why social topics amount to roughly
two thirds of human everyday communication \citep{dunbar1997human}
and
why mind-wandering and dreams
are so rich in stories about people and
the complex relationships between them
\citep{schilbach2008minds}.
%
Consequently, the DMN has been proposed to reflect
continuous processing of social information as a
physiological baseline of human brain function
\citep{schilbach2008minds}. This view was later challenged by observing
analogues of the DMN in monkeys \citep{mantini2011default}
and rats \citep{lu2012rat}, species with
supposedly less advanced social-cognitive capacities.
Instead,
our formal account can parsimoneously explain the dominance of social content in
human mental scenes by their relevance to humans
reflected in high estimates for social information obtained from the value function
\citep{baker2009action}.
In sum,
our account on the DMN directly implies
its previously proposed ``radar'' function
of monitoring the environment for salient information
in general and inherently incorporates the importance of social environmental cues
as an important special case, without excluding other topics
contemplated in continuous mental activity.
